---
title: "R Notebook"
output: html_notebook
---

#0. Load packages
```{r}
library(dplyr)
library(Seurat)
library(tictoc)
library(reticulate)
library(limma)#from Biocmanager
library(viridisLite)
library(patchwork)
library(future)
plan('multisession', workers = 5)#cores = 5
options(future.globals.maxSize = 100000 * 1024^2)

library(CellChat)
library(patchwork)
options(stringsAsFactors = FALSE)

```

#Subset the Seurat object
```{r}
##Subset
P21_cc <- subset(P21_Mouse, idents = c('0 DP', '1 Coch Fibro', '2 RF', '3 BV', 
                     '5 LBV', '6 BV', '8 PF', '9 Pericyte', 
                     '10 Krtno', '11 BV', '12 Krtno', '14 Krtno', 
                     '15 Fascia', '17 DS', '18 APM'), invert = FALSE)#subset the fibroblast supercluster
P21_cc#view the subset

##Rename clusters to their #'s since this cleans up the CellChat plots
cluster_names <- c('DP', '1', '2', '3', 
                     '5', '6', '8', '9', 
                     '10', '11', '12', '14', 
                     '15', '17', '18')
#cluster_names <- as.data.frame(strsplit2(x = cluster_names, split = ' '))

new.cluster.ids <- cluster_names
#new.cluster.ids[1] <- 'O'#cellchat can't have cluster ids that are 0
names(new.cluster.ids) <- levels(P21_cc)
P21_cc <- RenameIdents(P21_cc, new.cluster.ids)
```

#Initialize cellchat object and db
```{r}
data.input <- GetAssayData(P21_cc, assay = "SCT", slot = "data") # normalized data matrix
labels <- Idents(P21_cc)
meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels


cellchat <- createCellChat(object = data.input, meta = meta, group.by = "group")
#> Create a CellChat object from a data matrix
#> Set cell identities for the new CellChat object
#> The cell groups used for CellChat analysis are  APOE+ FIB FBN1+ FIB COL11A1+ FIB Inflam. FIB cDC1 cDC2 LC Inflam. DC TC Inflam. TC CD40LG+ TC NKT
#> 
#> cellchat <- addMeta(cellchat, meta = meta)
cellchat <- setIdent(cellchat, ident.use = "group") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels
groupSize <- as.numeric(table(cellchat@idents)) # number of cells in each cell group

#set Species database
CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)

# Show the structure of the database
dplyr::glimpse(CellChatDB$interaction)
#> Rows: 1,939
#> Columns: 11
#> $ interaction_name   <chr> "TGFB1_TGFBR1_TGFBR2", "TGFB2_TGFBR1_TGFBR2", "TGFB…
#> $ pathway_name       <chr> "TGFb", "TGFb", "TGFb", "TGFb", "TGFb", "TGFb", "TG…
#> $ ligand             <chr> "TGFB1", "TGFB2", "TGFB3", "TGFB1", "TGFB1", "TGFB2…
#> $ receptor           <chr> "TGFbR1_R2", "TGFbR1_R2", "TGFbR1_R2", "ACVR1B_TGFb…
#> $ agonist            <chr> "TGFb agonist", "TGFb agonist", "TGFb agonist", "TG…
#> $ antagonist         <chr> "TGFb antagonist", "TGFb antagonist", "TGFb antagon…
#> $ co_A_receptor      <chr> "", "", "", "", "", "", "", "", "", "", "", "", "",…
#> $ co_I_receptor      <chr> "TGFb inhibition receptor", "TGFb inhibition recept…
#> $ evidence           <chr> "KEGG: hsa04350", "KEGG: hsa04350", "KEGG: hsa04350…
#> $ annotation         <chr> "Secreted Signaling", "Secreted Signaling", "Secret…
#> $ interaction_name_2 <chr> "TGFB1 - (TGFBR1+TGFBR2)", "TGFB2 - (TGFBR1+TGFBR2)…

```

#Select the database to use downstream
```{r}
# use a subset of CellChatDB for cell-cell communication analysis
#CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling") # use just Secreted Signaling
# use all CellChatDB for cell-cell communication analysis
CellChatDB.use <- CellChatDB

# set the used database in the object
cellchat@DB <- CellChatDB.use
#__________________________________________________________________________________________________

# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database

```

#Compute the communication probability and infer cellular communication network
```{r}
tic("Identifying overexpressed genes and interactions")
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)#population.size = TRUE?

toc()
# project gene expression data onto PPI (Optional: when running it, USER should set `raw.use = FALSE` in the function `computeCommunProb()` in order to use the projected data)
# cellchat <- projectData(cellchat, PPI.human)
```

#Calculate Interactions At the Community Level (Longest Step)
```{r fig.height=7, fig.width=7}
tic("Computing Communities")

cellchat <- computeCommunProb(cellchat)
# Filter out the cell-cell communication if there are only few number of cells in certain cell groups
cellchat <- filterCommunication(cellchat, min.cells = 3)#we have some small clusters I want to keep

cellchat <- computeCommunProbPathway(cellchat)
toc()


cellchat <- aggregateNet(cellchat)
```


```{r}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")

```

Save Cellchat object now
```{r}
save(cellchat, file = '/t2tb/Object/Phan2020_P21_Mouse_cellchat.Rdata')
```

#If loading from saved CellChat object:
```{r}
#groupSize <- as.numeric(table(cellchat@idents)) # number of cells in each cell group
```



#View CirclePlot of overall interactions between clusters
```{r fig.height=10, fig.width=10}
#____
mat <- cellchat@net$weight

path_out <- '~/Desktop/Sean_CellChat/Plots/P21_EpiDer/Cluster_CirclePlot/'
setwd(path_out)#need to change working directory to export plots b/c the plot is not a ggplot object
umap_cols <- c('#00441b', '#238b45', '#74c476', '#8c6bb1', '#ffab4b',
                     '#88419d', '#7850a7', '#fc8d59', '#006d2c', '#7fcdbb', 
                     '#4d8ad5', '#a58bc2', '#00206a', '#f16913', '#16539e', 
                     '#a1d99b', '#800026', '#41ab5d', '#f768a1', '#a50f15')#string vector of my custom colors for UMAP
cols_i <- as.numeric(levels(P21_cc))
cols_i[1] <- 0#since cellchat can't take 0 as a cluster id, now need to replace 'chr_id' --> NA from as.numeric w/ 0 for i
umap_cols <- umap_cols[cols_i + 1]


par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i], color.use = umap_cols)
  
  #use png graphics device to export the cellchat plots (gives control over dimensions and naming versus netVisual function)
  png(filename = paste0(rownames(mat)[i], ' Interactions.png', sep = ''),
      width = 7, height = 7, units = 'in', res = 300)#start graphics device
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i], color.use = umap_cols)
  dev.off()
}

#_____cellchat@netP$pathways "MK" "PTN""PERIOSTIN" "SPP1" "ANGPTL""EGF""PDGF" "MIF"GAS""VEGF""GRN""SEMA3""VISFATIN""TNF""FGF"CXCL"PROS""WNT"
#___"ANGPT" "TGFb""BMP" "ncWNT" "CCL""CALCR""KIT"     "CHEMERIN"  "CSF"       "NPY"       "HGF"       "NGF"       "NT"      

```

#View CirclePlot of overall pathway interactions
```{r fig.height=5, fig.width=7}
pathways.show <- cellchat@netP$pathways#store pathways with significant communication
pathways.show#print to notebook the pathways

# Hierarchy plot
# Here we define `vertex.receive` so that the left portion of the hierarchy plot shows signaling to fibroblast and the right portion shows signaling to immune cells 
vertex.receiver = seq(1,4) # a numeric vector.
path_out <- '~/Desktop/Sean_CellChat/Plots/P21_EpiDer/Path_CirclePlot/'
setwd(path_out)#need to change working directory to export plots b/c the plot is not a ggplot object

n_loop <- 1#storage variable to iterate through
while (n_loop <= length(pathways.show)) {
  netVisual_aggregate(cellchat, signaling = pathways.show[n_loop],  vertex.receiver = vertex.receiver,
                      show.legend = FALSE, thresh = 0.01, color.use = umap_cols)#want to view one pathway at a time
  
  #use png graphics device to export the cellchat plots (gives control over dimensions and naming versus netVisual function)
  png(filename = paste0(pathways.show[n_loop], '.png', sep = ''),
      width = 7, height = 7, units = 'in', res = 300)#start graphics device
  netVisual_aggregate(cellchat, signaling = pathways.show[n_loop],  vertex.receiver = vertex.receiver,
                      show.legend = FALSE, thresh = 0.05, color.use = umap_cols)#want to view one pathway at a time
  dev.off()
  
  n_loop <- n_loop + 1#iterate to next pathway
}


```


#Netvisual aggregate & heatmap
```{r fig.height=7, fig.width=7}
# Circle plot
par(mfrow=c(1,1))
netVisual_aggregate(cellchat, signaling = pathways.show, layout = "circle")

# Chord diagram
par(mfrow=c(1,1))
netVisual_aggregate(cellchat, signaling = pathways.show, layout = "chord")

# Heatmap
par(mfrow=c(1,1))
n_loop <- 1#storage variable to iterate through
while (n_loop <= length(pathways.show[1:length(pathways.show)])) {
  #print(pathways.show[n_loop])#test
  print(netVisual_heatmap(cellchat, signaling = pathways.show[n_loop], color.heatmap = "Reds"))#want to view one pathway at a time
  n_loop <- n_loop + 1#iterate to next pathway
}

n_loop <- 46#storage variable to iterate through; 45 errors so skipping
while (n_loop <= length(pathways.show[46:length(pathways.show)])) {
  #print(pathways.show[n_loop])#test
  print(netVisual_heatmap(cellchat, signaling = pathways.show[n_loop], color.heatmap = "Reds"))#want to view one pathway at a time
  n_loop <- n_loop + 1#iterate to next pathway
}

#> Do heatmap based on a single object
#netVisual_circle(cellchat@netP$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

#Compute the contribution of each ligand-receptor pair to the overall signaling pathway and visualize cell-cell communication mediated by a single ligand-receptor pair
```{r}
pathways.show <- cellchat@netP$pathways#store pathways with significant communication
pathways.show#print to notebook the pathways

#view contribution of each pathways's ligand receptor pairs
n_loop <- 1#storage variable to iterate through
while (n_loop <= length(pathways.show)) {
  p1 <- netAnalysis_contribution(cellchat, signaling = pathways.show[n_loop]) + 
    ggtitle(pathways.show[n_loop])#want to view one pathway at a time
  print(p1)
  
  n_loop <- n_loop + 1#iterate to next pathway
}

#netAnalysis_contribution(cellchat, signaling = pathways.show)#view all pathways
```

#View circle plots of ligand receptor (LR) pairs
Vignette does not do this automated so I wrote a nested loop to print circle plots for each ligand receptor pair within each pathway
```{r fig.height=7, fig.width=7}
pathways.show <- cellchat@netP$pathways#store pathways with significant communication

path_out <- '~/Desktop/Sean_CellChat/Plots/P21_EpiDer/LR_Pair_CirclePlot/'
setwd(path_out)#need to change working directory to export plots b/c the plot is not a ggplot object

n_pathway <- 1#storage variable to iterate through pathways
while (n_pathway <= length(pathways.show)) {
  LR.temp <- extractEnrichedLR(cellchat, signaling = pathways.show[n_pathway], 
                               geneLR.return = FALSE, thresh = 0.05,                      
                               enriched.only = TRUE)#return sig. LR pairs from a single pathway
  
  n_lr <- 1#storage variable to iterate through ligand receptor pairs within a pathway
  while (n_lr <= nrow(LR.temp)) {
    print(paste(pathways.show[n_pathway], LR.temp[n_lr, ], sep = ': '))
    vertex.receiver = seq(1, (nrow(LR.temp) + 1)) # a numeric vector
    p1 <- netVisual_individual(cellchat, signaling = pathways.show[n_pathway],  
                         pairLR.use = LR.temp[n_lr, ], vertex.receiver = vertex.receiver,
                         arrow.size = 0.3, arrow.width = 1.5, color.use = umap_cols)
    
      #use png graphics device to export the cellchat plots (gives control over dimensions and naming versus netVisual function)
    png(filename = paste0(pathways.show[n_pathway], '-', LR.temp[n_lr, ], '.png', sep = ''),
        width = 7, height = 7, units = 'in', res = 300)#start graphics device
    netVisual_individual(cellchat, signaling = pathways.show[n_pathway],  
                         pairLR.use = LR.temp[n_lr, ], vertex.receiver = vertex.receiver,
                         arrow.size = 0.3, arrow.width = 1.5, color.use = umap_cols)
    dev.off()
    
    n_lr <- n_lr + 1#iterate to next ligand receptor pair within the pathway
  }
  
  n_pathway <- n_pathway + 1#iterate to next pathway
}

```

